### 2024-08-26 chapters/bci.tex
Add wcble paper

\todo{chapters don't nicely follow imrad structure. Is this  problem?}
\todo{concern: is a bit in between formula 1 and formula 2}
\section{Rationale}
\section{Overview \& state-of-the-art}
different recording technologies and level of invasiveness + infographic
active/reactive/passive + infographic
\section{Visual Event-Related Potential based BCIs}

>>>>

\todo{chapters don't nicely follow imrad structure. Is this  problem?}
\todo{concern: is a bit in between formula 1 and formula 2}
\section{Rationale}
\todo{people helped and societal impact medical vs recreational}
\todo{figure BCI loop}
\todo{definition bci society}
\section{Overview \& state-of-the-art}
\subsection{Recording technologies}
\subsection{BCI paradigms}
\subsection{Decoders}
~\cite{Review of classification techniques}
different recording technologies and level of invasiveness + infographic
active/reactive/passive + infographic
\section{Visual Event-Related Potential based BCIs}

### 2024-08-26 chapters/wcble.tex
Add wcble paper

The WCBLE algorithm is presented in Algorithm~\ref{alg:wcble}.
Its training phase is visualized in~Figure~\ref{fig:diagram-train}.
The initial training epochs $\{\mat{X}_n^{(1)}\}_{n=1}^N$ are set to $\{\mat{X}_n^\mathrm{train}\}_{n=1}^N$.
At every iteration, classifier $\mathcal{C}$ is trained like in CBLE:
\begin{equation}
  \theta^{(i)} =
  \mathrm{train}_\mathcal{C}(\{\mat{X}^{(i-1)}_n[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
\end{equation}
Next, latency $s_{l_n}^{(i)}$ is determined for every epoch $\mat{X}^{(i)}$ corresponding
to its class label $l_n$ using equation~\ref{eq:prob-latency}.
Finally, the training epochs $\mat{X}^{(i+1)}$ for the next iteration are determined by aligning
each original training epoch to the latency $s_{l_n}^{(i)}$ corresponding to its respective class
label.
\begin{equation}
  \mat{X}^{(i+1)}_n = \mathrm{align}(\mat{X}_n^\mathrm{train}, s_{l_n}^{(i)}) \quad \forall \quad n=1,\ldots,N
\end{equation}
Aligning is performed by shifting and zero-padding the signal to the right if
the latency is negative relative to the time window onset, and to the left if
positive, by the difference between the latency and the window onset.
\begin{figure*}
	\includegraphics[width=\linewidth]{figures/wcble/figure1.pdf}
	\caption{Schematic representation of the Woody Classifier Based Latency
    Estimation training phase.
    (1) The first-stage spatiotemporal binary classifier is trained on a set of
    epochs.
    (2) It is	then applied to time-shifted copies of the epochs to obtain scores
    and class probabilities	over time.
    (3) The medians of these probability distributions are assumed the class
    latencies.
    (4) The epochs are aligned to their corresponding	class latencies by
    shifting in time such that all latencies fall at the same moment.
    (5) The spatiotemporal classifier is then retrained on the aligned epochs
    for a	next iteration.
    (6) After the iterative process halts, the scores and	latencies obtained
    from the last iteration are used to train the second-stage
		classifier.}
	\label{fig:diagram-train}
\end{figure*}
The process halts after a fixed amount of iterations or when the estimated set
of latencies has been encountered before, indicating it ended up in a loop.
In the end, the procedure should result in enhanced classifier parameters $\theta^*$,
closer to those when there would be no jitter between epochs.
Note that using the median approach for robust latency estimation results in a
smoother yet longer convergence process compared to the maximum/minimum approach.
We can then apply the classifier with enhanced parameters $\theta^*$ in a CBLE
manner to unseen epochs as illustrated in Figure~\ref{fig:diagram-eval} to obtain a
vector of scores over time as in~Section~\ref{sec:cble} and the estimated latencies as
in~Section~\ref{sec:robust-latency}.
\begin{figure*}
	\includegraphics[width=\linewidth]{figures/wcble/figure2.pdf}
	\caption{Schematic representation of the Woody Classifier Based Latency
    Estimation test phase. (1) The
		first-stage	spatiotemporal binary classifier obtained from the training phase is
		applied to time-shifted copies of the epochs to obtain scores and class probabilities
		over time. (2) The medians of these probability distributions are assumed
		as the new class latencies. (3) The scores and class latencies are input to the
		trained second-stage classifier, which predicts the label of the epochs.}
	\label{fig:diagram-eval}
\end{figure*}
\section{Algorithms}

\begin{algorithm}[H]
	\textsc{Train}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mbfX_n^\mathrm{train}\}_{n=1}^N,
  \mathbf{l}^\mathrm{train}, \mathcal{C}(\cdot,f, \Pr), s_1, s_2$
  \begin{algorithmic}[1]
		\State $\theta \gets
			\mathrm{train}_\mathcal{C}(\{\mbfX_n^\mathrm{train}[:,s_1:s_2]\}_{n=1}^N,\mathbf{l})$
      \algorithmiccomment{Train stage 1}
		\For{$n=1\ldots N$}
      \algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{n,s}^\mathrm{train} \gets f(\mbfX_n^\mathrm{train}[:,s:s+(s_2-s_1)],\theta)$
        \EndFor
        \State $s_{n,\mathrm{target}}^\mathrm{train} \gets
			    \mathrm{median}\left[Pr(s|\mbfX_n^\mathrm{train},\theta,\mathrm{target})\right]$
		    \State $s_{n,\mathrm{non-target}}^\mathrm{train}\gets
			    \mathrm{median}\left[Pr(s|\mbfX_n^\mathrm{train},\theta,\textrm{non-target})\right]$
    \EndFor
	\end{algorithmic}
  \textbf{Output:} $\theta,
  \mbfY^\mathrm{train},
  \mathbf{s}_\mathrm{target}^\mathrm{train},
  \mathbf{s}_\mathrm{non-target}^\mathrm{train}$
	\smallskip \hrule \smallskip
	\textsc{Evaluate}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mbfX_m^\mathrm{test}\}_{m=1}^N, \mathcal{C}(\theta,f, \Pr), s_1, s_2$
  \begin{algorithmic}[1]
		\For{$m=1\ldots M$}
      \algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{m,s}^\mathrm{test} \gets f(\mbfX_n^\mathrm{test}[:,s:s+(s_2-s_1)],\theta)$
        \EndFor
        \State $s_{m,\mathrm{target}}^\mathrm{test} \gets
			    \mathrm{median}\left[Pr(s|\mbfX_m^\mathrm{test},\theta,\mathrm{target})\right]$
		    \State $s_{m,\mathrm{non-target}}^\mathrm{test}\gets
			    \mathrm{median}\left[Pr(s|\mbfX_m^\mathrm{test},\theta,\textrm{non-target})\right]$
    \EndFor
	\end{algorithmic}
  \textbf{Output:} $\mbfY^\mathrm{test},
  \mathbf{s}_\mathrm{target}^\mathrm{test},
  \mathbf{s}_\mathrm{non-target}^\mathrm{test}$

@@

\newpage
\begin{algorithm}[H]
	\textsc{Train}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mbfX_n^\mathrm{train}\}_{n=1}^N, \mathbf{l},
		\mathcal{C}(\cdot,f,\Pr), s_1, s_2$
	\begin{algorithmic}[1]
		\State $\mbfX'_n \gets \mbfX^\mathrm{train}_n \quad \forall \quad n=1\ldots
    N$
      \algorithmiccomment{Train stage 1}
		\Repeat
		\State $\theta^* \gets \mathrm{train}_\mathcal{C}(\{\mbfX'_n[:,s_1:s_2]\}_0^{N-1},\mathbf{l}$
		\For{$n=1\ldots N$}
    \State  $s_n \gets
			    \mathrm{median}\left[Pr(s|\mbfX_n',\theta^*,l_n)\right]$
		\State $\mbfX'_n \gets \mathrm{align}(\mbfX^\mathrm{train}_n, s^*_n)$
		\EndFor
    \Until{convergence or maximum iterations reached}
    \For{$n=1\ldots N$}\algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{n,s}^\mathrm{train} \gets
      f(\mbfX_n^\mathrm{train}[:,s:s+(s_2-s_1)],\theta^*)$
        \EndFor
        \State $s_{n,\mathrm{target}}^\mathrm{train} \gets
			    \mathrm{median}\left[Pr(s|\mbfX_n^\mathrm{train},\theta^*,\mathrm{target})\right]$
		    \State $s_{n,\mathrm{non-target}}^\mathrm{train}\gets
			    \mathrm{median}\left[Pr(s|\mbfX_n^\mathrm{train},\theta^*,\textrm{non-target})\right]$
    \EndFor

	\end{algorithmic}
	\textbf{Output:} $\theta^*,
  \mbfY^\mathrm{train},
  \mathbf{s}_\mathrm{target}^\mathrm{train},
  \mathbf{s}_\mathrm{non-target}^\mathrm{train}$

	\smallskip \hrule \smallskip
	\textsc{Evaluate}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mbfX_m^\mathrm{test}\}_{m=1}^N, \mathcal{C}(\theta^*,f, \Pr), s_1, s_2$
  \begin{algorithmic}[1]
		\For{$m=1\ldots M$}\algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{m,s}^\mathrm{test} \gets
      f(\mbfX_n^\mathrm{test}[:,s:s+(s_2-s_1)],\theta^*)$
        \EndFor
        \State $s_{m,\mathrm{target}}^\mathrm{test} \gets
			    \mathrm{median}\left[Pr(s|\mbfX_m^\mathrm{test},\theta^*,\mathrm{target})\right]$
		    \State $s_{m,\mathrm{non-target}}^\mathrm{test}\gets
			    \mathrm{median}\left[Pr(s|\mbfX_m^\mathrm{test},\theta^*,\textrm{non-target})\right]$
    \EndFor
	\end{algorithmic}
  \textbf{Output:} $\mbfY^\mathrm{test},
  \mathbf{s}_\mathrm{target}^\mathrm{test},
  \mathbf{s}_\mathrm{non-target}^\mathrm{test}$
	\caption{Classifier-Based Latency Estimation with Woody Iterations}
	\label{alg:wcble}
\end{algorithm}

>>>>

The WCBLE algorithm is presented in Algorithm~\ref{alg:wcble}.
Its training phase is visualized in~Figure~\ref{fig:diagram-train}.
The initial training epochs $\{\mat{X}_n^{(1)}\}_{n=1}^N$ are set to $\{\mat{X}_n^\mathrm{train}\}_{n=1}^N$.
At every iteration, classifier $\mathcal{C}$ is trained like in CBLE:
\begin{equation}
  \theta^{(i)} =
  \mathrm{train}_\mathcal{C}(\{\mat{X}^{(i-1)}_n[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
\end{equation}
Next, latency $s_{l_n}^{(i)}$ is determined for every epoch $\mat{X}^{(i)}$ corresponding
to its class label $l_n$ using equation~\ref{eq:prob-latency}.
Finally, the training epochs $\mat{X}^{(i+1)}$ for the next iteration are determined by aligning
each original training epoch to the latency $s_{l_n}^{(i)}$ corresponding to its respective class
label.
\begin{equation}
  \mat{X}^{(i+1)}_n = \mathrm{align}(\mat{X}_n^\mathrm{train}, s_{l_n}^{(i)}) \quad \forall \quad n=1,\ldots,N
\end{equation}
Aligning is performed by shifting and zero-padding the signal to the right if
the latency is negative relative to the time window onset, and to the left if
positive, by the difference between the latency and the window onset.
\begin{figure*}
	\includegraphics[width=\linewidth]{figures/wcble/figure1.pdf}
	\caption{Schematic representation of the Woody Classifier Based Latency
    Estimation training phase.
    (1) The first-stage spatiotemporal binary classifier is trained on a set of
    epochs.
    (2) It is	then applied to time-shifted copies of the epochs to obtain scores
    and class probabilities	over time.
    (3) The medians of these probability distributions are assumed the class
    latencies.
    (4) The epochs are aligned to their corresponding	class latencies by
    shifting in time such that all latencies fall at the same moment.
    (5) The spatiotemporal classifier is then retrained on the aligned epochs
    for a	next iteration.
    (6) After the iterative process halts, the scores and	latencies obtained
    from the last iteration are used to train the second-stage
		classifier.}
	\label{fig:diagram-train}
\end{figure*}
The process halts after a fixed amount of iterations or when the estimated set
of latencies has been encountered before, indicating it ended up in a loop.
In the end, the procedure should result in enhanced classifier parameters $\theta^*$,
closer to those when there would be no jitter between epochs.
Note that using the median approach for robust latency estimation results in a
smoother yet longer convergence process compared to the maximum/minimum approach.
We can then apply the classifier with enhanced parameters $\theta^*$ in a CBLE
manner to unseen epochs as illustrated in Figure~\ref{fig:diagram-eval} to obtain a
vector of scores over time as in~Section~\ref{sec:cble} and the estimated latencies as
in~Section~\ref{sec:robust-latency}.
\begin{figure*}
	\includegraphics[width=\linewidth]{figures/wcble/figure2.pdf}
	\caption{Schematic representation of the Woody Classifier Based Latency
    Estimation test phase. (1) The
		first-stage	spatiotemporal binary classifier obtained from the training phase is
		applied to time-shifted copies of the epochs to obtain scores and class probabilities
		over time. (2) The medians of these probability distributions are assumed
		as the new class latencies. (3) The scores and class latencies are input to the
		trained second-stage classifier, which predicts the label of the epochs.}
	\label{fig:diagram-eval}
\end{figure*}
\section{Algorithms}
\todo{insert comments}
\begin{algorithm}[H]
	\textsc{Train}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mat{X}_n^\mathrm{train}\}_{n=1}^N,
  \mathbf{l}^\mathrm{train}, \mathcal{C}(\cdot,f, \Pr), s_1, s_2$
  \begin{algorithmic}[1]
		\State $\theta \gets
			\mathrm{train}_\mathcal{C}(\{\mat{X}_n^\mathrm{train}[:,s_1:s_2]\}_{n=1}^N,\mathbf{l})$
      %\algorithmiccomment{Train stage 1}
		\For{$n=1\ldots N$}
      %\algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{n,s}^\mathrm{train} \gets f(\mat{X}_n^\mathrm{train}[:,s:s+(s_2-s_1)],\theta)$
        \EndFor
        \State $s_{n,\mathrm{target}}^\mathrm{train} \gets
			    \mathrm{median}\left[Pr(s|\mat{X}_n^\mathrm{train},\theta,\mathrm{target})\right]$
		    \State $s_{n,\mathrm{non-target}}^\mathrm{train}\gets
			    \mathrm{median}\left[Pr(s|\mat{X}_n^\mathrm{train},\theta,\textrm{non-target})\right]$
    \EndFor
	\end{algorithmic}
  \textbf{Output:} $\theta,
  \mat{Y}^\mathrm{train},
  \mathbf{s}_\mathrm{target}^\mathrm{train},
  \mathbf{s}_\mathrm{non-target}^\mathrm{train}$
	\smallskip \hrule \smallskip
	\textsc{Evaluate}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mat{X}_m^\mathrm{test}\}_{m=1}^N, \mathcal{C}(\theta,f, \Pr), s_1, s_2$
  \begin{algorithmic}[1]
		\For{$m=1\ldots M$}
      %\algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{m,s}^\mathrm{test} \gets f(\mat{X}_n^\mathrm{test}[:,s:s+(s_2-s_1)],\theta)$
        \EndFor
        \State $s_{m,\mathrm{target}}^\mathrm{test} \gets
			    \mathrm{median}\left[Pr(s|\mat{X}_m^\mathrm{test},\theta,\mathrm{target})\right]$
		    \State $s_{m,\mathrm{non-target}}^\mathrm{test}\gets
			    \mathrm{median}\left[Pr(s|\mat{X}_m^\mathrm{test},\theta,\textrm{non-target})\right]$
    \EndFor
	\end{algorithmic}
  \textbf{Output:} $\mat{Y}^\mathrm{test},
  \mathbf{s}_\mathrm{target}^\mathrm{test},
  \mathbf{s}_\mathrm{non-target}^\mathrm{test}$

@@

\newpage
\begin{algorithm}[H]
	\textsc{Train}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mat{X}_n^\mathrm{train}\}_{n=1}^N, \mathbf{l},
		\mathcal{C}(\cdot,f,\Pr), s_1, s_2$
	\begin{algorithmic}[1]
		\State $\mat{X}'_n \gets \mat{X}^\mathrm{train}_n \quad \forall \quad n=1\ldots
    N$
      %\algorithmiccomment{Train stage 1}
		\Repeat
		\State $\theta^* \gets \mathrm{train}_\mathcal{C}(\{\mat{X}'_n[:,s_1:s_2]\}_0^{N-1},\mathbf{l}$
		\For{$n=1\ldots N$}
    \State  $s_n \gets
			    \mathrm{median}\left[Pr(s|\mat{X}_n',\theta^*,l_n)\right]$
		\State $\mat{X}'_n \gets \mathrm{align}(\mat{X}^\mathrm{train}_n, s^*_n)$
		\EndFor
    \Until{convergence or maximum iterations reached}
    \For{$n=1\ldots N$}%\algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{n,s}^\mathrm{train} \gets
      f(\mat{X}_n^\mathrm{train}[:,s:s+(s_2-s_1)],\theta^*)$
        \EndFor
        \State $s_{n,\mathrm{target}}^\mathrm{train} \gets
			    \mathrm{median}\left[Pr(s|\mat{X}_n^\mathrm{train},\theta^*,\mathrm{target})\right]$
		    \State $s_{n,\mathrm{non-target}}^\mathrm{train}\gets
			    \mathrm{median}\left[Pr(s|\mat{X}_n^\mathrm{train},\theta^*,\textrm{non-target})\right]$
    \EndFor

	\end{algorithmic}
	\textbf{Output:} $\theta^*,
  \mat{Y}^\mathrm{train},
  \mathbf{s}_\mathrm{target}^\mathrm{train},
  \mathbf{s}_\mathrm{non-target}^\mathrm{train}$

	\smallskip \hrule \smallskip
	\textsc{Evaluate}
	\smallskip \hrule \smallskip
	\textbf{Input:} $\{\mat{X}_m^\mathrm{test}\}_{m=1}^N, \mathcal{C}(\theta^*,f, \Pr), s_1, s_2$
  \begin{algorithmic}[1]
		\For{$m=1\ldots M$}%\algorithmiccomment{Feature extraction for stage 2}
		  \For{$s=1\ldots R$}
      \State $y_{m,s}^\mathrm{test} \gets
      f(\mat{X}_n^\mathrm{test}[:,s:s+(s_2-s_1)],\theta^*)$
        \EndFor
        \State $s_{m,\mathrm{target}}^\mathrm{test} \gets
			    \mathrm{median}\left[Pr(s|\mat{X}_m^\mathrm{test},\theta^*,\mathrm{target})\right]$
		    \State $s_{m,\mathrm{non-target}}^\mathrm{test}\gets
			    \mathrm{median}\left[Pr(s|\mat{X}_m^\mathrm{test},\theta^*,\textrm{non-target})\right]$
    \EndFor
	\end{algorithmic}
  \textbf{Output:} $\mat{Y}^\mathrm{test},
  \mathbf{s}_\mathrm{target}^\mathrm{test},
  \mathbf{s}_\mathrm{non-target}^\mathrm{test}$
	\caption{Classifier-Based Latency Estimation with Woody Iterations}
	\label{alg:wcble}
\end{algorithm}
